---
title: "GIS Wk4 Homework Task"
author: "J Post"
date: "2022-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(terra)
library(sf)
library(tidyverse)
library(here)
library(countrycode)
library(tmap)
library(tmaptools)
```

## Task

Download gender inequality index (GII) data from the UNDP, and join to spatial data of the World. Create a new column of difference in inequality between 2010 and 2019. Plot the map, and then share on Github.

GII data is downloaded from [here](https://hdr.undp.org/data-center/documentation-and-downloads) (Table 5)  
World spatial data is downloaded from [here](https://hub.arcgis.com/datasets/esri::world-countries-generalized/about)

### Load in Data

```{r loaddata}
genderindex <- read_csv(
  here("Data", "HDR21-22_Composite_indices_complete_time_series.csv")
  # Think about the 'na =' and 'locale' arguments
  #   e.g. review the csv first to see what the na is represented as
  )

worldshp <- read_sf(here("Data", "World_Countries__Generalized_.shp"))

```

## Clean and format

- Select the columns for years 2010 and 2019
- Calculate new column for change in index between years
- Convert country ID to a shared code format
- Join to shapefile data

```{r cleandata}
gendercln <- genderindex %>% 
  # Remember to use 'clean_names' (janitor package)
  select(iso3:region, gii_2010, gii_2019) %>%
  mutate(giidiff = gii_2019 - gii_2010,
         giidiff_pc = gii_2010/(gii_2019-gii_2010),
         iso2 = countrycode(iso3, 
                            origin = "iso3c", 
                            destination = "iso2c")
  )

genderjoin <- worldshp %>%
  left_join(gendercln, by = c("ISO" = "iso2"))

```

## Plot the map of GII difference

```{r plotmap}

tmap_mode("plot")

giimap <- tm_shape(genderjoin) +
  tm_polygons("giidiff",
              style = "jenks",
              palette = "PRGn") +
  tm_layout(main.title = "Change in Gender Inequality Index, 2010 to 2019",
            main.title.position = "center")
giimap

# **NOTE
#   See Andy's script from Wk4/5 homework task repo for code on faceted maps
#    and formatting a shared legend and caption

```

Create faceted map with shared legend

```{r plot_facetmap}

breaks_index <- c(0, 0.20, 0.35, 0.45, 0.6, 0.8)
breaks_diff <- c(-0.5, -0.25, -0.08, -0.05, -0.02, 0.1)

map_gii2010 <- tm_shape(genderjoin) +
  tm_polygons("gii_2010",
              breaks = breaks_index,
              # style = "jenks",
              palette = "PuRd") +
  tm_layout(frame = FALSE) + 
  tm_legend(show = FALSE) +
  tm_credits("(a) GII by Country, 2010", position = c(0,0.9), size = 0.8)

map_gii2019 <- tm_shape(genderjoin) +
  tm_polygons("gii_2019",
              breaks = breaks_index,
              # style = "jenks",
              palette = "PuRd") +
  tm_layout(frame = FALSE) + 
  tm_legend(show = FALSE) +
  tm_credits("(b) GII by Country, 2019", position = c(0,0.9), size = 0.8)

map_giidiff <- tm_shape(genderjoin) +
  tm_polygons("giidiff",
              breaks = breaks_diff,
              # style = "jenks",
              palette = "PRGn") +
  tm_layout(frame = FALSE) + 
  tm_legend(show = FALSE) +
  tm_credits("(c) Change in GII, 2010 to 2019", position = c(0,0.9), size = 0.8)

legend <- tm_shape(genderjoin) +
  tm_polygons("gii_2010",
              breaks = breaks_index,
              # style = "jenks",
              palette = "PuRd") +
  tm_layout(legend.only = TRUE, legend.position=c(0.2,0.25), asp=0.1,
            title = "Gender Inequality Index") +
  tm_credits("Source: United Nations Development Program (UNDP)")

facetmap <- tmap_arrange(map_gii2010, map_gii2019, map_giidiff, legend)
facetmap

```


